---
title: "Penalized Logistic Regression"
author: "Nick Wornson"
date: "1/6/2021"
output: html_document
---

```{r setup, include=FALSE,message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pROC)
library(MASS)
library(glmnet)
library(caret)
source('utilities.r')
```


# Load the data

Data is scaled and imputed in preprocessing.ipynb.  Test data is than scaled using the same scale and imputation values as the training data.  load_data() and train_val_split() are custom functions located in 'utilities.r'
```{r}
data = load_data()
split = train_val_split(data$data,.7)
train = split$train
val = split$val

h1_col = which(colnames(train) == 'h1n1_vaccine')
seas_col = which(colnames(train) == 'seasonal_vaccine')

h1_train = train[,-seas_col]
h1_val = val[,-seas_col]

s_train = train[,-h1_col]
s_val = val[,-h1_col]

```


# Data preprocessing.  

glmnet requires us to create the design matrix for input
```{r}
# model matrices required for glmnet
MMH = model.matrix(h1n1_vaccine ~ .,h1_train)[,-1] # remove the intercept
MMS = model.matrix(seasonal_vaccine ~ .,s_train)[,-1]

MMHV = model.matrix(h1n1_vaccine ~ .,h1_val)[,-1]
MMSV = model.matrix(seasonal_vaccine ~ .,s_val)[,-1]

```

# Try out a lasso regression model

We can start with a lasso regression, and perform cross-validation to get an optimal lambda value.  

```{r}
# find the best alpha using lasso regression

cv.lasso.h = cv.glmnet(MMH,h1_train$h1n1_vaccine,
                     alpha = 1,
                     family = 'binomial')

cv.lasso.h$lambda.min
plot(cv.lasso.h)
```

Lets fit the model using the optimal lambda value and look at the AUC score on the validation set
```{r message=FALSE}
opto.model.h = glmnet(MMH,h1_train$h1n1_vaccine,
                    alpha = 1, 
                    family = "binomial",
                    lambda = cv.lasso.h$lambda.min)

cv.lasso.probs.h = opto.model.h %>% 
  predict(newx = MMHV,type = 'response')
score=auc(as.numeric(h1_val$h1n1_vaccine),as.numeric(cv.lasso.probs.h))
```


Now lets create a loop to go through each combination of lasso, ridge, and elasticnet for each response variable, find the optimal lambda value for each, and report the auc on the validation set.  I have created a function to perform the cv, fit the model, and return the AUC score located in the utilities.r file.  


```{r message=FALSE}

results = data.frame(response = character(),
                     score = numeric(),
                     alpha = numeric(),
                     lambda = numeric())

alphas = c(0,.2,.4,.5,.6,.7,.8,1)
for(i in 1:length(alphas)){
  
  h1_score = cv.lasso.best(MMH,h1_train$h1n1_vaccine,MMHV,
                           h1_val$h1n1_vaccine,alphas[i])
  
  h1_out = data.frame(response = 'h1n1',
                      score = h1_score$score,
                      alpha = alphas[i],
                      lambda = h1_score$lambda)
  
  s_score = cv.lasso.best(MMS,s_train$seasonal_vaccine,
                          MMSV,s_val$seasonal_vaccine,alphas[i])
  
  s_out = data.frame(response = 'seasonal',
                      score = s_score$score,
                      alpha = alphas[i],
                      lambda = s_score$lambda)
  
  results = rbind(results,h1_out,s_out)
  
  

  
}

```


```{r pressure, echo=FALSE}
results$score = as.numeric(results$score)
results
```
Not much distinction between the models.. 

# Fit our optimal models
```{r}
best = results %>% group_by(response) %>%
    filter(score == max(score))
best
```
# Make a submission
```{r}
alpha.h = best[1,3]
lambda.h = best[1,4]
alpha.s = best[2,3]
lambda.s = best[2,4]

MMH_full = model.matrix(h1n1_vaccine ~ .,data$data[,-seas_col])[,-1] 
MMS_full = model.matrix(seasonal_vaccine ~ .,data$data[,-h1_col])[,-1]

MMtest = model.matrix( ~ .,data$test)[,-1]

opto.model.h = glmnet(MMH_full,data$data$h1n1_vaccine,
                    alpha = alpha.h, 
                    family = "binomial",
                    lambda = lambda.h)

opto.model.s = glmnet(MMS_full,data$data$seasonal_vaccine,
                    alpha = alpha.s, 
                    family = "binomial",
                    lambda = lambda.s)

cv.elastic.probs.h = opto.model.h %>% 
    predict(newx = MMtest,type = 'response')

cv.elastic.probs.s = opto.model.s %>% 
    predict(newx = MMtest,type = 'response')

submit = data.frame(respondent_id = data$test_id,
                    h1n1_vaccine = cv.elastic.probs.h,
                    seasonal_vaccine = cv.elastic.probs.s)
colnames(submit) = c('respondent_id','h1n1_vaccine','seasonal_vaccine')

write.csv(submit,'elastic_cv.csv',row.names = FALSE)

# scored .8350

```


```{r}
head(submit)
```


